--
title: "Chapter 8"
output: 
  html_document:
     code_folding: show
     theme: flatly
     highlight: tango
     toc: yes
     number_sections: yes
     df_print: paged
     toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
ed_df <- read.csv('c:/Statistical_Methods/education_career_success.csv')
ed_df

```
First we analyze the data.

In this section we will investigate the number of alumni that went on to have senior level jobs, this will be our success, and 
mid and entry level will count as our failure. In this section we will take a simple random sample of 50 alumni to determine the number of success is from binomial distribution.

# Section 1.1

```{r}
# Senior Job level will be treated as success
ed_df %>%
  count(Current_Job_Level)

prop <- 59/(80+261)
prop
```

##a)
1. n = 70, fixed number.
2. Trials are independent and do not affect the outcome of other.
3. There are only 2 outcomes, Senior = success, all others = failure.
4. p = 0.02174981, is constant from the given population data.


##b)
Now we will find the mean of the of the binomial random sample, so that we can determine the probability of the random varaible
is more than its mean.

```{r}
mean_sample <- 70 * prop
mean_sample
```
We will check the probability that the random variable is more than its mean

```{r}
1 - pbinom(q=mean_sample, size=70, prob=prop)

```
We find that the probability that the random variable is more than 8.651026 to be  0.4377059.

##c)
Next we describe the sampling distribution.

population = 400 alumni
sample = 70 random alumni
variable = Alumni in senior level positions
p_hat = .1730205
sd =.05349475
10.016 > 10

```{r}

n <- 70
p <- prop

mean_phat <- p
sd_phat <- sqrt(p * (1 - p) / n)

mean_phat
sd_phat

70*mean_phat*(1-mean_phat)

```




## Section 1.2
The quantitative variable for this section is Salary. We will first check the normal distribution of the variable.

Population: 

```{r}
ed_df %>%
  ggplot(mapping = aes(x = Starting_Salary)) +
  geom_histogram(fill="salmon")

ed_df %>%
  ggplot(mapping = aes(sample = Starting_Salary)) +
  stat_qq() +
  stat_qq_line(color="red")

# population mean 
mu <- mean(ed_df$Starting_Salary)

# population std dev
sigma <- sd(ed_df$Starting_Salary)


#Creating sample of 20
ed_sample_df <- sample_n(ed_df, size = 20,replace = F)

```
X = IQ
population = all alumni
sample = 20 alumni
sample size = 20 *(n=20)* alumni
Xbar = mean starting salary of 20 random alumni
$\mu x = mean starting salary of all alumni$
$\mu bar{x}$ = mean of sample starting salary of 20 random alumni for all samples

# Introduction



##1.




# Simple Linear Regression
In this section we are going to investigate if there is a linear relationship on whether a higher GPA reflects a higher starting salary.

```{r}
ed_lm <- lm(formula = Starting_Salary ~ University_GPA, data = ed_df)
ed_lm


ed_df %>% 
  ggplot(mapping=aes(x=Starting_Salary, y=University_GPA)) +
  geom_point() +
  geom_smooth(method="lm",color="red", se=F) 

```
