--
title: "Chapter 8"
output: 
  html_document:
     code_folding: show
     theme: flatly
     highlight: tango
     toc: yes
     number_sections: yes
     df_print: paged
     toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```




```{r}
ed_df <- read.csv('c:/Statistical_Methods/education_career_success.csv')
ed_df

```
First we analyze the data.

In this section we will investigate the number of alumni that went on to have senior level jobs, this will be our success, and 
mid and entry level will count as our failure. In this section we will take a simple random sample of 50 alumni to determine the number of success is from binomial distribution.

# Probability Distributions

## Section 1.1

```{r}
# Senior Job level will be treated as success
ed_df %>%
  count(Current_Job_Level)

prop <- 59/400
prop


```

###a)
1. n = 70, fixed number.
2. Trials are independent and do not affect the outcome of other.
3. There are only 2 outcomes, Senior = success, all others = failure.
4. p = 0.1475, is constant from the given population data.


###b)
Now we will find the mean of the of the binomial random sample, so that we can determine the probability of the random varaible
is more than its mean.

```{r}
mean_sample <- 70 * prop
mean_sample
```
We will check the probability that the random variable is more than its mean

```{r}
1 - pbinom(q=mean_sample, size=70, prob=prop)

```
We find that the probability that the random variable is more than 10.325 to be 0.4606525.

###c)
Next we describe the sampling distribution.

population = 400 alumni
sample = 70 random alumni
variable = Alumni in senior level positions
p_hat = .1730205
sd =.05349475
10.016 > 10

```{r}

n <- 70
p <- prop

mean_phat <- p
sd_phat <- sqrt(p * (1 - p) / n)

mean_phat
sd_phat

70*mean_phat*(1-mean_phat)

```


## Section 1.2
The quantitative variable for this section is Career_Satisfaction. We will first check the normal distribution of the variable.

Population: 

```{r}
ed_df %>%
  ggplot(mapping = aes(x = Career_Satisfaction)) +
  geom_histogram(fill="salmon")

ed_df %>%
  ggplot(mapping = aes(sample = Career_Satisfaction)) +
  stat_qq() +
  stat_qq_line(color="red")


```
X = Career Satisfaction Report
population = all alumni
sample = 20 alumni
sample size = 20 *(n=20)* alumni
$\bar{x} = 7.8$ = mean career satisfaction of 20 random alumni 
$\mu x = 7.7925$  mean career satisfaction of all alumni

Histogram Plot shows a skew to the left, but qq plot looks approximatly normal.

We will find the probability of career satisfaction being greater than 8 for a random selected student.
P(X>8)

```{r}
# population mean 
mu <- mean(ed_df$Career_Satisfaction)
mu
# population std dev
sigma <- sd(ed_df$Career_Satisfaction)


#Creating sample of 20
ed_sample_df <- sample_n(ed_df, size = 20,replace = F)
#ed_sample_df

x_bar <- mean(ed_sample_df$Career_Satisfaction)
x_bar

1 - pnorm(q=8, mean=mu, sd=sigma/sqrt(20))

```
P(X>8) = .2495

# Introduction

##1.
In this analysis we are exploring the links between performance in school linked to career success and satisfaction after school.
This is important to understand that the effort that students put in to school work now will benefit them long after school is over with. 

a)

The following are research questions in this analysis:

1. What is the proportion of alumni who achieve senior-level positions, and what is a confidence interval estimate for this proportion?

2. What is the average career satisfaction of alumni, and what is a confidence interval estimate for the population mean?

3. Is the proportion of alumni in senior-level positions greater than 15%?

4. Is the average career satisfaction of alumni different from 7.12?

5. Is there a difference in average starting salaries between male and female alumni?

6. Is there a linear relationship between university GPA and career satisfaction?

These questions will be addressed using statistical methods including confidence intervals, hypothesis testing, and regression analysis.


##2.
We, generally, spend 40, sometimes more, hours per week at our careers during the week. Getting into the right career that we are happy or at least content with is important for overall well-being and mental health. This data analysis aims to tie together the importance of doing our best in school, because it is easy to slack off, because it could indeed shape our future selves sense of well-being.

##3.
We are using the Education & Career Success datset on Kaggle. This dataset explores how students' academic achievements, skills, and extracurricular activities translate into real-world career outcomes.

Link to our dataset:
https://www.kaggle.com/datasets/adilshamim8/education-and-career-success

The dataset has 400 alumni recorded from various disciplines.The dataset gathers information in 3 key sections that have several different columns. The sections are Academic Performance, Skills & Extracurriculars, and Career Outcomes. Within these sections we have between 4 - 7 columns each to have enough data to draw conclusions from. Our groups variables of interest are, University_GPA, Career_Satisfaction, Current_Job_Level, and ... **NEEED MORE INFO**




# Exploratory Data Analysis

## Section 3.0

a)Quantitative:
Career Satisfaction:
```{r}
ed_df %>% 
  summarize(mean_sat = mean(Career_Satisfaction, na.rm = TRUE), median_sat = median(Career_Satisfaction, na.rm = TRUE))
range(ed_df$Career_Satisfaction)
ed_df %>% 
  ggplot(mapping = aes(x = Career_Satisfaction)) + 
  geom_histogram(color = "black", fill = "pink", binwidth = 1) +
  labs(title = "Distribution of Alumni Career Satisfaction", 
       x = "Career Satisfaction Score (1-10)", 
       y = "Frequency (Count)")
```

Center: Mean = 7.7925, Median = 8
Shape: Skewed-left
Spread:Range = 5-10

Salary:
```{r}
ed_df %>% 
  summarize(mean_sal = mean(Starting_Salary, na.rm = TRUE), median_sal = median(Starting_Salary, na.rm = TRUE))
range(ed_df$Starting_Salary)
ed_df %>% 
  ggplot(mapping = aes(x = Starting_Salary)) + 
  geom_histogram(color = "black", fill = "pink") +labs(title = "Distribution of Alumni Starting Salaries",
       x = "Starting Salary ($)",
       y = "Count of Alumni")
```

Center:Mean = $87,562.50, Median = $84,000
Shape: Unimodal
Spread:Range = $31,000-$152,000

University GPA
```{r}
ed_df %>% 
  summarize(mean_gpa = mean(University_GPA, na.rm = TRUE), median_gpa = median(University_GPA, na.rm = TRUE))
range(ed_df$University_GPA)
ed_df %>% 
  ggplot(mapping = aes(x = University_GPA)) + 
  geom_histogram(color = "black", fill = "pink", binwidth = 0.1) +
  labs(title = "Distribution of University GPA",
       x = "University GPA",
       y = "Count of Alumni")
```

Center: Mean = 3.441, Median = 3.45
Shape: Approximately normal
Spread: Range = 2.8-4.0

b) Qualitative
Current Job Level:
```{r}
ed_df %>% 
  count(Current_Job_Level, name = "freq") 

ed_df %>% 
  ggplot(aes(x = Current_Job_Level)) + 
  geom_bar(fill = "pink", color = "black") + 
  labs(title = "Current Job Levels", x = "Job Level", y = "Count of Alumni")
```

Gender:
```{r}
ed_df %>% 
  count(Gender, name = "freq") 

ed_df %>% 
  ggplot(aes(x = Gender)) + 
  geom_bar(fill = "pink", color = "black") + 
  labs(title = "Gender Distribution", x = "Gender", y = "Count of Alumni")
```



# One Sample: Confidence Interval and Hypothesis Testing

## Section 4.0: Population and Sample

a) Population: All alumni
b) Sample: 400 alumni

## Section 4.1: Quantitative Variables

 
a) Variable: Career Satisfaction of alumni - Quantitative
b) Parameter: Mean career satisfaction of all alumni
c) Statistic: Mean career satisfaction of 400 alumni

```{r}
# Point Estimate
ed_df %>%
  summarise(mean_sal = mean(Career_Satisfaction),
            sd(x=Career_Satisfaction))

x_bar <- 7.7925
x_bar_sd <- 1.372645

BSDA::tsum.test(mean.x = x_bar, s.x = x_bar_sd, n.x=400,
                conf.level = .99)


```
X_bar = 7.7925

Decision: We are 99% confident that the mean career satisfaction for alumni is between 7.614866 and 7.970134.

e) We find that the mean career satisfaction of an alumni is 7.12 at a signigicance level of 0.05. Can we expect the career satisfaction to be different than the mean?


$H_0: \mu = 7.12$
$H_A: \mu \ne 7.12$

* n $\le$ 0.05N (independent samples) 
* Simple random sample 
* 400 > 30

```{r}
BSDA::tsum.test(mean.x = 7.7925, 
                s.x = 1.372645, 
                n.x = 400,
                mu = 7.12,
                alternative = "two.sided")

```
p-value: 2.2e-16
Decision: Since p-value = .0000 < alpha 0.05, We reject $H_0$.
Conclusion: We reject $H_0$, There is sufficient evidence at $\alpha = 0.05$ to say that the average career satisfaction of alumni is not equal to  7.12.



## Section 4.2: Qualitative Variables

a) Variable: Alumni in senior level positions - Qualtitative
b) Parameter: Proportion of all alumni in senior level positions
c) Statistic: Proportion of 400 alumni in senior level positions

```{r}
ed_df %>% count(Current_Job_Level)
proportion <- 59/ 400
proportion

prop.test(x=59, n=400, conf.level = .98)
```
d)
Point Estimate = .1475
We are 98% confident that the proportion of alumni in senior level positions is between 0.1098738 and 0.1948188.

e) We find that on average 15% of alumni end up in senior level positions at a signigicance level of 0.05. Can we expect more alumni in senior level positions?

__Step 1__: Form the hypothesis.

$\H_0: p = .15$ 
$\H_A: p \gt .15$ 

__Step 2__: What is the level of significance, $\alpha$?

$\alpha = 0.05$


__Step 0__: Check requirement

* n $\le$ 0.05N (independent samples) 
* Simple random sample 
* $51 \gt 10 $

```{r}
# checking requirement
400*.15*(1-.15)

prop.test(x=59, n=400, p=.15, alternative = "greater")
```
Decision: Since p-value = 0.5279 > 0.05 alpha. We do not reject the null hypothesis.

Conclusion: We do not reject the null hypothesis. There is insufficient evidence at alpha = 0.05 to say that the proportion of alumni in senior level positions is greater than .15.



# Two Samples t-test - Independent Sampling Method

## Section 5.0
###a)
What are the populations? 
Population 1: All male Students
Population 2: All female students
###b)
What are the samples?
sample 1: 200 male students
sample 2: 200 female students
###c)
What are the variables? What type of variables are they?
starting salary - quantitative
###d)
What are the parameters?
$\mu_m$ - mean starting salaries of all male students
$\mu_f$ - mean starting salaries of all female students
###e)
What are the statistics?
$\bar{x}_m$ - mean starting salaries of 200 male students
$\bar{x}_f$ - mean starting salaries of 200 female students
###f)
Run a hypothesis test to test your hypothesis.
Requirements:
SRS
$n_1 = 200 < 0.05N_1 and n_2 = 200 < 0.05N_2$
$n_1 = 200 > 30 and n_2 = 200 > 30$
Hypothesis:
$H_0:\mu_m - \mu_f = 0$
$H_A:\mu_m - \mu_f \neq 0$
let $\alpha = 0.05$
```{r}
male_students <- ed_df %>% filter(Gender == "Male")
female_students <- ed_df %>% filter(Gender == "Female")
t.test(x = male_students$Starting_Salary, y = female_students$Starting_Salary, mu = 0, alternative = "two.sided")

```
p-value = 0.0008928
Decision: Since the p-value = 0.0008928 < $\alpha = 0.05$, we reject $H_0$.
Conclusion: There is sufficient evidence to suggest that there is a starting salary difference between the male and female students. 
###g)
Use the Confidence interval to test your hypothesis.
```{r}
t.test(x = male_students$Starting_Salary, y = female_students$Starting_Salary, conf.level = 0.99)
```
Conclusion: We are 99% confident that the mean starting salary difference between male and female students is between -17402.975 and -2227.025. Since 0 is not included in the confidence interval and all of the interval is negative, we reject $H_0$.There is sufficient evidence that female graduates will earn more on average than male graduates at their first job post graduation.



# Simple Linear Regression
In this section we are going to investigate if there is a linear relationship on whether students that work harder to achieve a higher GPA reflects a higher reporting of career satisfaction with a significance level of 0.05.

$H_0: \beta_1 = 0$
$H_A: \beta_1 \ne 0$

$\alpha = 0.05$

```{r}
ed_lm <- lm(formula = Career_Satisfaction ~ University_GPA, data = ed_df)
ed_lm


ed_df %>% 
  ggplot(mapping=aes(x=Career_Satisfaction, y=University_GPA)) +
  geom_point() +
  geom_smooth(method="lm",color="red", se=F) 

ed_lm %>% summary()
```
P-value: .0000
Test Statistic: 75.62

Y-intercept Interpretation: Since GPA cannot be 0, there is no meaningful interpretation of the y-intercept.

Decision: Since P-value = .0000 < alpha = 0/05, we reject $H_0$.

Conclusion: We reject $H_0$. There is sufficient evidence at alpha = 0.05, to say that there is a linear relationship between a high GPA and later having more satisfaction in their career.

