---
title: "Math 3311 Chapter 1"
author: "Josh Pippen"
output:
  html_document:
      df_print: paged
      highlight: tango
      theme: flatly
      toc: yes
      toc_depth: 2
      toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "c:/Statistical_Methods/chapter1")
```

https://rmarkdown.rstudio.com/articles_docx.html

Loading packages
```{r warning=F, message=F}
library(tidyverse)
```
------------------------------------------------------------------------------
------------------------------------------------------------------------------


## Section 1.2 Pictorial and Tabular Methods in Descriptive Statistics

-------------------------------------------------------------------------------

### Example 5

The following categories for type of physical activity involved when an industrial accident occurred appeared in the article “Finding Occupational Accident Patterns in the Extractive Industry Using a Systematic Data Mining Approach” (Reliability Engr. and System Safety, 2012: 108–122):
A. Working with handheld tools
B. Movement
C. Carrying by hand
D. Handling of objects
E. Operating a machine
F. Others
```{r warning=F, message=F}
accidents_df <- read_csv("ex01-29.csv")
accidents_df
```

a) Create a frequency distribution of the accident data.
```{r}
#Check the class of the object
accidents_df %>% class()

#Frequency Table
accidents_df %>% count(Accidents, name = "freq")

```

b) Create a relative frequency distribution of the accident data.
```{r}
# Relative Frequency
accidents_df %>% 
  count(Accidents, name = "freq") %>%
  mutate(rel_freq = freq/sum(freq))

```

c) Create a bar graph of the accident data.
```{r}
# Using raw data
accidents_df %>%
  ggplot(mapping = aes(x = Accidents)) +
  geom_bar(fill = "coral", color="black") +
  labs(y = "Frequency")

    
```

d) Comment on the distribution of the accident data.


-------------------------------------------------------------------------------


### Example 6

The accompanying data set consists of observations on shower-flow rate (L/min) for a sample of n = 129 houses in Perth, Australia (“An Application of Bayes Methodology to the Analysis of Diary Records in a Water Use Study,” J. Amer. Stat. Assoc., 1987: 705–711)

```{r warning=F, message=F}
#rate_df <- read_csv("ex01-14.csv")
#rate_df
#rate_v <- rate_df %>% pull(Rate)
```

a) Create a frequency distribution of the shower-flow rate data.
```{r}


```

b) Create a relative frequency distribution of the shower-flow rate data.
```{r}


```

c) Create a histogram of the shower-flow rate data.
```{r}


```

d) Comment on the distribution of the shower-flow rate data.


-------------------------------------------------------------------------------
-------------------------------------------------------------------------------


## Section 1.3 Measure of Location

### Example 7

The accompanying data set consists of observations on shower-flow rate (L/min) for a sample of n = 129 houses in Perth, Australia (“An Application of Bayes Methodology to the Analysis of Diary Records in a Water Use Study,” J. Amer. Stat. Assoc., 1987: 705–711)
Calculate the following:

a) Mean, median, mode
```{r}


```


b) Q1, Q3, 90th percentile
```{r}


```


c) Create boxplot and violin plot. 
```{r}


```


d) $\bar{x}_{tr(10)}$
```{r}


```

-------------------------------------------------------------------------------

-------------------------------------------------------------------------------


## Section 1.4 Measure of Variability

### Example 7

The accompanying data set consists of observations on shower-flow rate (L/min) for a sample of n = 129 houses in Perth, Australia (“An Application of Bayes Methodology to the Analysis of Diary Records in a Water Use Study,” J. Amer. Stat. Assoc., 1987: 705–711)
Calculate the following:
a) Range, standard deviation, variance, IQR
```{r}


```

-------------------------------------------------------------------------------
-------------------------------------------------------------------------------

